Visualizing the Geography of Genetic Variants

### 2016-07-01 paper.Rmd
filled in more captions

# Abstract {-}
One of the key characteristics of any genetic variant, beyond its potential phenotypic effects or its frequency, is its geographic distribution. The geographic distribution of a genetic variant can shed light on where the variant first arose, in what populations it spread within, and in turn can help one learn about historical patterns of migration, genetic drift, and natural selection. Collectively the geographic distribution of genetic variants can help to explain how populations have been related through time (e.g. levels of gene flow and divergence). For these reasons, visual inspection of geographic maps for genetic variants is common practice in genetic studies. Here we develop an interactive visualization tool for illuminating the geographic distribution of genetic variants. Through an efficient REST API and dynamic front-end the Geography of Genetic Variants (GGV) browser rapidly provides maps of allele frequencies in populations distributed across the globe.

@@

```{r figure1, echo=FALSE, message=FALSE, fig.height=5, fig.width=5, fig.align='center', fig.cap='Schematic diagram of the Geography of Genetic Variants (GGV) web application. We compute minor allele frequencies from publicly available population genetic datasets from the 1000 Genomes Project, Human Genom Diversity Project (HGDP) and Population Reference Sample (POPRES). These allele freqeuncies are stored in a SQL database along side with geographic coordinates associated with each population. The front-end visualization was created using d3 and accesses allele frequency and population meta-data from a REST API implemented in the python library flask.'}
img <- readPNG("figures/png/figure1.png")
grid.raster(img)
```

@@

```{r figure2, echo=FALSE, message=FALSE, fig.height=5, fig.width=5, fig.align='center', fig.cap='Example map from the Geography of Genetic Variants web application. Each pie chart represents a population with the blue slice of the pie displaying the freqeuncy of the global minor allele and the yellow slice of the pie displaying the frequency of the global major allele in each population.'}
img <- readPNG("figures/png/figure2.png")
grid.raster(img)
```

@@

```{r figure3, echo=FALSE, message=FALSE, fig.height=6, fig.width=6, fig.align='center', fig.cap='Figure caption'}
img <- readPNG("figures/png/figure3.png")
grid.raster(img)
```

@@

```{r figure4, echo=FALSE, message=FALSE, fig.height=6, fig.width=6, fig.align='center', fig.cap='Figure caption'}
img <- readPNG("figures/png/figure4.png")
grid.raster(img)
```

@@

```{r figure5, echo=FALSE, message=FALSE, fig.height=6, fig.width=6, fig.align='center', fig.cap='Figure caption'}
img <- readPNG("figures/png/figure5.png")
grid.raster(img)
```

>>>>

# Abstract {-}
One of the key characteristics of any genetic variant, beyond its potential phenotypic effects or its frequency, is its geographic distribution. The geographic distribution of a genetic variant can shed light on where the variant first arose, in what populations it spread within, and in turn can help one learn about historical patterns of migration, genetic drift, and natural selection. Collectively the geographic distribution of genetic variants can help to explain how populations have been related through time (e.g. levels of gene flow and divergence). For these reasons, visual inspection of geographic maps for genetic variants is common practice in genetic studies. Here we develop an interactive visualization tool for illuminating the geographic distribution of genetic variants. Through an efficient REST API and dynamic front-end the [\textit{Geography of Genetic Variants (GGV)}](popgen.uchicago.edu/ggv) browser rapidly provides maps of allele frequencies in populations distributed across the globe.

@@

```{r figure1, echo=FALSE, message=FALSE, fig.height=5, fig.width=5, fig.align='center', fig.cap='Schematic diagram of the Geography of Genetic Variants (GGV) web application. We compute minor allele frequencies from publicly available population genetic datasets from the 1000 Genomes Project, Human Genome Diversity Project (HGDP) and Population Reference Sample (POPRES). These allele frequencies are stored in a SQL database along side with geographic coordinates associated with each population. The front-end visualization was created using d3 and accesses allele frequency and population meta-data from a REST API implemented in the python library flask.'}
img <- readPNG("figures/png/figure1.png")
grid.raster(img)
```

@@

```{r figure2, echo=FALSE, message=FALSE, fig.height=5, fig.width=5, fig.align='center', fig.cap='Example map from the Geography of Genetic Variants web application. Each pie chart represents a population with the blue slice of the pie displaying the frequency of the global minor allele and the yellow slice of the pie displaying the frequency of the global major allele in each population.'}
img <- readPNG("figures/png/figure2.png")
grid.raster(img)
```

@@

```{r figure3, echo=FALSE, message=FALSE, fig.height=6, fig.width=6, fig.align='center', fig.cap='Example map from the Geography of Genetic Variants web application displaying the use of varying transparency of population pie charts to represent uncertainty in allele frequencies. The transparency is scaled in proportion to the number of observed chromosomes for a particular variant in each population.'}
img <- readPNG("figures/png/figure3.png")
grid.raster(img)
```

@@

```{r figure4, echo=FALSE, message=FALSE, fig.height=6, fig.width=6, fig.align='center', fig.cap='Example map from the Geography of Genetic Variants web application displaying the use of frequency scales for more expressive representations of rare variation on geographic maps. The blue pie charts convey a give minor allele frequency out of 100 percent, the green out 1 percent, and the red out of .01 percent.'}
img <- readPNG("figures/png/figure4.png")
grid.raster(img)
```

@@

```{r figure5, echo=FALSE, message=FALSE, fig.height=6, fig.width=6, fig.align='center', fig.cap='Example map from the Geography of Genetic Variants web application displaying the use of a force directed layout to limit visual clutter when many populations overlap in geographic position. The left map shows the original population locations while the right shows the application of the force directed.'}
img <- readPNG("figures/png/figure5.png")
grid.raster(img)
```

### 2016-07-01 paper.tex
filled in more captions

One of the key characteristics of any genetic variant, beyond its
potential phenotypic effects or its frequency, is its geographic
distribution. The geographic distribution of a genetic variant can shed
light on where the variant first arose, in what populations it spread
within, and in turn can help one learn about historical patterns of
migration, genetic drift, and natural selection. Collectively the
geographic distribution of genetic variants can help to explain how
populations have been related through time (e.g.~levels of gene flow and
divergence). For these reasons, visual inspection of geographic maps for
genetic variants is common practice in genetic studies. Here we develop
an interactive visualization tool for illuminating the geographic
distribution of genetic variants. Through an efficient REST API and
dynamic front-end the Geography of Genetic Variants (GGV) browser
rapidly provides maps of allele frequencies in populations distributed
across the globe.

@@

\caption[Schematic diagram of the Geography of Genetic Variants (GGV) web application]{Schematic diagram of the Geography of Genetic Variants (GGV) web application. We compute minor allele frequencies from publicly available population genetic datasets from the 1000 Genomes Project, Human Genom Diversity Project (HGDP) and Population Reference Sample (POPRES). These allele freqeuncies are stored in a SQL database along side with geographic coordinates associated with each population. The front-end visualization was created using d3 and accesses allele frequency and population meta-data from a REST API implemented in the python library flask.}\label{fig:figure1}
\end{figure}
\end{CodeChunk}

@@

\caption[Example map from the Geography of Genetic Variants web application]{Example map from the Geography of Genetic Variants web application. Each pie chart represents a population with the blue slice of the pie displaying the freqeuncy of the global minor allele and the yellow slice of the pie displaying the frequency of the global major allele in each population.}\label{fig:figure2}
\end{figure}
\end{CodeChunk}

@@

\caption[Figure caption]{Figure caption}\label{fig:figure3}
\end{figure}
\end{CodeChunk}

@@

\caption[Figure caption]{Figure caption}\label{fig:figure4}
\end{figure}
\end{CodeChunk}

@@

\caption[Figure caption]{Figure caption}\label{fig:figure5}
\end{figure}
\end{CodeChunk}

>>>>

One of the key characteristics of any genetic variant, beyond its
potential phenotypic effects or its frequency, is its geographic
distribution. The geographic distribution of a genetic variant can shed
light on where the variant first arose, in what populations it spread
within, and in turn can help one learn about historical patterns of
migration, genetic drift, and natural selection. Collectively the
geographic distribution of genetic variants can help to explain how
populations have been related through time (e.g.~levels of gene flow and
divergence). For these reasons, visual inspection of geographic maps for
genetic variants is common practice in genetic studies. Here we develop
an interactive visualization tool for illuminating the geographic
distribution of genetic variants. Through an efficient REST API and
dynamic front-end the
\href{popgen.uchicago.edu/ggv}{\textit{Geography of Genetic Variants (GGV)}}
browser rapidly provides maps of allele frequencies in populations
distributed across the globe.

@@

\caption[Schematic diagram of the Geography of Genetic Variants (GGV) web application]{Schematic diagram of the Geography of Genetic Variants (GGV) web application. We compute minor allele frequencies from publicly available population genetic datasets from the 1000 Genomes Project, Human Genome Diversity Project (HGDP) and Population Reference Sample (POPRES). These allele frequencies are stored in a SQL database along side with geographic coordinates associated with each population. The front-end visualization was created using d3 and accesses allele frequency and population meta-data from a REST API implemented in the python library flask.}\label{fig:figure1}
\end{figure}
\end{CodeChunk}

@@

\caption[Example map from the Geography of Genetic Variants web application]{Example map from the Geography of Genetic Variants web application. Each pie chart represents a population with the blue slice of the pie displaying the frequency of the global minor allele and the yellow slice of the pie displaying the frequency of the global major allele in each population.}\label{fig:figure2}
\end{figure}
\end{CodeChunk}

@@

\caption[Example map from the Geography of Genetic Variants web application displaying the use of varying transparency of population pie charts to represent uncertainty in allele frequencies]{Example map from the Geography of Genetic Variants web application displaying the use of varying transparency of population pie charts to represent uncertainty in allele frequencies. The transparency is scaled in proportion to the number of observed chromosomes for a particular variant in each population.}\label{fig:figure3}
\end{figure}
\end{CodeChunk}

@@

\caption[Example map from the Geography of Genetic Variants web application displaying the use of frequency scales for more expressive representations of rare variation on geographic maps]{Example map from the Geography of Genetic Variants web application displaying the use of frequency scales for more expressive representations of rare variation on geographic maps. The blue pie charts convey a give minor allele frequency out of 100 percent, the green out 1 percent, and the red out of .01 percent.}\label{fig:figure4}
\end{figure}
\end{CodeChunk}

@@

\caption[Example map from the Geography of Genetic Variants web application displaying the use of a force directed layout to limit visual clutter when many populations overlap in geographic position]{Example map from the Geography of Genetic Variants web application displaying the use of a force directed layout to limit visual clutter when many populations overlap in geographic position. The left map shows the original population locations while the right shows the application of the force directed.}\label{fig:figure5}
\end{figure}
\end{CodeChunk}
